---
title: "Submission for MOD005246"
subtitle: "Assignment E010"
author: "SID: 205707"
date: "`r Sys.Date()`"
output: 
    word_document:
    pdf_document:
    html_document:
bibliography: ../Bibliography/references.bib
csl: ../Bibliography/harvard-anglia-ruskin-university.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse, quietly = TRUE)
```

# Popular science articles

## Build it ... or not, and they will come

Discussion of open-ended vs 'conventional' species targeted conservation - old wine in new bottles?

Nature in Britain is in trouble. We have lost half our biodiversity over the last few hundred years according to recent analysis led by the [Natural History Museum](https://www.theguardian.com/p/j72x7), making Britain one of the most nature-depleted regions in Europe (see figure).

```{r, echo=FALSE, results="hide"}
path <- here::here("data/")
#list.files(path)
bii <- read_csv(glue::glue(path, "resource.csv"),show_col_types = FALSE)
```

```{r, echo=FALSE}
bii %>%
  janitor::clean_names() %>%
  filter(str_detect(area_code, "154-") , 
         scenario == "historical", 
         variable == "bii") %>%
  mutate(Country = str_remove(area_code, "001-150-154-")) %>%## UN M49 codes for Northern Europe
  ggplot(aes(year, value, colour = Country)) +
  geom_line() +
  labs(title = "Biodiversity intactness in Northern European countries", 
       subtitle = "Britain lags behind", 
       y = "Biodiversity Intactness Index", 
       caption = "Source = Natural History Museum data portal") +
  theme_bw()
  

```

Using an indicator called the [Biodiversity Intactness Index (BII)](https://www.nhm.ac.uk/our-science/data/biodiversity-indicators/about-the-biodiversity-intactness-index.html) which calculates the percentage of species remaining and their frequency pre- and post- human impact in a given area, Britain is estimated to have only 50% of its pre-agrarian and industrial revolution biodiversity and continues a slow decline.

This loss is largely attributable to urbanisation and agricultural practices. The latter is nowhere more visible than in the Fens - a vast former wetland in the East of England - where only about 0.2% of the original undrained fen exists. Over the centuries the Fens have been systematically drained for agricultural land, which has degraded as top soil and peat have been lost.

At Wicken Fen - a tiny remnant of undrained fen in a vast sea of agricultural land - the impact is clear to see. The fen is some 2-3 metres higher than than the surrounding area and is a mixture of wetland, reedbedsm pasture, grassland and woodland which sustains enormous biodiversity - over 9000 species at last count.

Restoring fenland is high on the agenda - landscape fragments are not sustainable - so there is a plan to knit together existing undrained fen remnants and restore tracts of arable farmland nearing their end of life to create a corridor restored fenland from Wicken to Cambridge, increasing the size of the reserve from 255 to 53 0 hectares.

.As well as being a step to restoring habitat and biodiversity, restoring and rewetting also increases carbon capture and storage. It has been estimated that 100,000 tonnes of CO2 area released every year from peat loss on the Fens as a whole.

Rewetting and rewilding. Open-ended landscape restoration. Big vision. Need the land. Maintining the original fen (oldest nature reserve in Britain 1899) is a high intensity task involving cutting reed on a 3 year rotation, preventing the fen from returning to carr and woodland. Sedge and reed was cut by hand although recently the NT has acquired a reed cutting machine.

The restoration programme

\[Hectares\]



## Eager beavers

After an absence of 400 years beavers (Castor fiber) are back in Britain.  Hunted to extinction for fur, meat and secretions (castoreum), beavers disappeared from Britain by the end of the 16th century, but now they are being reintroduced across the UK.

Beavers are a 'keystone' species and are arch ecosystem engineers and this has been a prime motivation for reintroductions especially given the success of programmes across Europe, as well as increased interest in restoring native species.

Beavers are exclusive herbivores, eating bark, rhizomes, leaves and twigs, with a particular predilection for willow. A recent review by Brazier et al has identified myriad ways in which beavers impact their environment including:
•	Building dams which slow water flow and help control flooding, and build channels which help create wetlands
•	Resulting ponds provide habitat for plants and other species, and are nutrient rich and protect beavers from predators
•	Harvesting trees for dam building and food which has a 'coppicing' effect on understorey growth
•	Improving water quality
•	Creation of new habitat
Taken together the evidence strongly suggests that beavers provide a wide range of valuable  ecosystem services and increase biodiversity. As well as increasing biodiversity, beaver engineering of water systems may help mitigate against flooding caused by extreme rain events, and  generated pools and dams capture carbon and create carbon sinks through the accumulation of sediment and new plant growth.
.
Scottish Beaver Trial
This longest running reintroduction programme in Britain is the Scottish Beaver Trial which began in 2008 in Knapdale Forest with the release of 16 beavers from Norway. The project was inspired by the success of reintroduction programmes across Europe where the beaver had historically suffered a similar fate to that in Britain and was long in the gestation – first being mooted in the 1990s.

This programme has been extensively studied and evaluated and is the inspiration for 

  
Beaver and dam at Spain's Hall Estate, Essex.  Julian Flowers


River Otter Beaver Trial

In 2014, wild beavers were found living on the River Otter in Devon presumably having escaped from captivity. After careful work, DEFRA was persuaded to support a trial rather than recapturing them to return to captivity. This has been extensively studies culminating in 

2021 – the year of the beaver
There are now numerous proposed or live reintroductions across the country and 2021 has been a particularly active year.
•	Derbyshire Wildlife Trust – beavers were introduced on its Willington Wetlands Nature reserve in September 2021
•	Ham Fen, Kent where there are now 10 beavers following a release in 2001
•	Welsh Beaver project – spring 2021 release at the Cors Dyfi reserve in Montgomeryshire
•	Cornwall Beaver project – 2017 release near Truro
•	Cheshire – Hatchmere nature reserve in 2020
•	Dorset - 2 beavers released in 2021 in West Dorset
•	Cumbria – 2 beavers released in the Lake District in 2020
•	Essex – introduction of beavers as part of a natural flood management plan at Spain’s Hall Estate to protect the village of Finchingfield
In all it is estimated that there are now nearly 200 beavers in Britain.
Beavers and humans
Despite initial concerns beaver-human interactions seem largely positive. Reintroduction projects are creating eco-tourism and fundraising opportunities and engaging local people through media and social media campaigns.
Beavers have to introduced under licence (Schedule 9 of the Wildlfie and Countryside Act 1981).
A recent review by Natural England in lieu of a national consulationm concluded that across a wide range of habitat and species, beaver reintroduction was likely to promote increased diversity but for some groups more research was needed. There is potential for conflict where beavers create territories which impinge on low lying arable farms on flood plains but some of these issues are anticipated and addressed in the Beaver Management Handbook.

Its not all plain sailing – although beaver are required to be introduced into enclosed environments, they have proved adept at escaping and the Scotland experience demonstrates they can establish wild, free-living populations.

What next?

There are now several hundred beavers across the UK and DEFRA are now cosulting on beaver reintroduction in Englad

https://consult.defra.gov.uk/natural-environment-policy/beaver-reintroduction-and-management/ 

Beavers are back and they may be coming to a location near you…

